Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-08-24 grove-paper.tex
Added some TikZ figures showing graphs

\tikzset{graphs/graph style/.style={
 tree layout,
 level sep=1cm,
 nodes={draw, rounded corners=1mm},
 edges={draw, sloped, font={\small\ttfamily}}
}}
\tikzset{graphs/root style/.style={
 %draw=none,
 as={\textbullet$_0$}
}}
%\tikzset{graphs/every edge/.style={}

@@

@@

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
TODO
\caption{User interface}
\label{fig:A simple tree with a hole-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_2$" [> "root", > "1"'] -> {
   "$\texttt{var(x)}_4$" [> "left", > "3"'],
   {}
  }
 }
};
\end{tikzpicture}
\caption{Graph representation}
\label{fig:A simple tree with a hole-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{A simple tree with a hole}
\label{fig:A simple tree with a hole}
\end{figure}

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
TODO
\caption{User interface}
\label{fig:Addition-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_2$" [> "root", > "1"'] -> {
   "$\texttt{var(x)}_4$" [> "left", > "3"'],
   "$\texttt{var(y)}_6$" [> "right", > "5"']
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Addition-b} after adding \vertex{6} and \edge{5}}
\label{fig:Addition-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Addition}
\label{fig:Addition}
\end{figure}

@@

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_2$" [> "root", > "1"'] -> {
   "$\texttt{var(x)}_4$" [> "left", > "3"'],
   {}
  }
 }
};
\path (2cm,-2cm) graph[graph style] {
 y [as={$\texttt{var(y)}_6$}]
};
\end{tikzpicture}
\caption{\autoref{fig:Addition-b} after deleting \edge{5}}
\label{fig:Change-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_2$" [> "root", > "1"'] -> {
   "$\texttt{var(x)}_4$" [> "left", > "3"'],
   "$\texttt{var(z)}_8$" [> "right", > "7"']
  }
 }
};
\path (2cm,-2cm) graph[graph style] {
 "$\texttt{var(y)}_6$"
};
\end{tikzpicture}
\caption{\autoref{fig:Change-a} after adding \vertex{8} and \edge{7}}
\label{fig:Change-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Change}
\label{fig:Change}
\end{figure}

@@

\begin{figure}
% TODO: naive add that creates a conflict
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
% \path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
% \path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
% \path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
% \path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style]
};
\path (-1cm,-3cm) graph[graph style] {
 "$\texttt{*}_2$" -> {
  "$\texttt{var(x)}_4$" [> "left", > "3"'],
  "$\texttt{var(z)}_6$" [> "right", > "5"']
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Change-b} after deleting \edge{1}}
\label{fig:Wrapping-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_8$" [> "root", > "7"']
 }
};
\path (-1cm,-3cm) graph[graph style] {
 "$\texttt{*}_2$" -> {
  "$\texttt{var(x)}_4$" [> "left", > "3"'],
  "$\texttt{var(z)}_6$" [> "right", > "5"']
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Wrapping-a} after adding \vertex{8} and \edge{7}}
\label{fig:Wrapping-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_8$" [> "root", > "7"'] -> {
   "$\texttt{*}_2$" [> "left", > "9"'] -> {
    "$\texttt{var(x)}_4$" [> "left", > "3"'],
    "$\texttt{var(z)}_6$" [> "right", > "5"']
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Wrapping-b} after adding \edge{9}}
\label{fig:Wrapping-c}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Wrapping}
\label{fig:Wrapping}
\end{figure}

@@

\begin{figure}
% \centering
% \begin{tikzpicture}
% \path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
% \path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
% \path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_8$" [> "root", > "7"']
 }
};
\path (0cm,-3cm) graph [graph style] {
 "$\texttt{*}_2$" -> {
  "$\texttt{var(x)}_4$" [> "left", > "3"'],
  "$\texttt{var(z)}_6$" [> "right", > "5"']
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Wrapping-c} after deleting \edge{9}}
\label{fig:Move-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_8$" [> "root", > "7"'] -> {
   {},
   "$\texttt{*}_2$" [> "right", > "11"'] -> {
    "$\texttt{var(x)}_4$" [> "left", > "3"'],
    "$\texttt{var(z)}_6$" [> "right", > "5"']
   }
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Move-a} after adding \edge{11}}
\label{fig:Move-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Moving \vertex{2} from the left child to the right child of \vertex{8}}
\label{fig:Move}
\end{figure}

@@

@@

\begin{figure}
\centering
TODO: add screenshot of code text to each node
TODO: remove orphans from diagram
\begin{tikzpicture}
\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/1.dot}};
\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2A.dot}};
\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2B.dot}};
\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/3.dot}};
\end{tikzpicture}
\caption{Editing Different Parts of the Code}
\label{fig:Editing Different Parts of the Code}
\end{figure}

@@

\begin{figure}
\centering
TODO: add screenshot of code text to each node
TODO: remove orphans from diagram
TODO: annotate edges with delete and restore
\begin{tikzpicture}
\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/1.dot}};
\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2A.dot}};
\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2B.dot}};
\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/3.dot}};
\end{tikzpicture}
\caption{Editing Nested Parts of the Code}
\label{fig:Editing Nested Parts of the Code}
\end{figure}

@@

\begin{figure}
\centering
\begin{tikzpicture}
\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/1.dot}};
\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2A.dot}};
\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2B.dot}};
\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/3.dot}};
\end{tikzpicture}
\caption{Direct Conflicts}
\label{fig:Direct Conflicts}
\end{figure}

Finally, we address genuine conflicts between edits.
Consider the top graph of \autoref{fig:Direct Conflicts}.
Suppose Alice inserts the number literal~\texttt{2} into the body of the lambda
while Bob inserts the function application~\texttt{f x} also into the body of the lambda.
These results in the left and right graphs of \autoref{fig:Direct Conflicts} respectively.
In particular pay attention to the edges going out of the \texttt{body} index of \vertex{3}.
In the left graph it is \edge{10}, which points to a number literal,
and in the right graph it is \edge{13}, which points to a function application.

We merge these by including \emph{both} edges as edges going out of the \texttt{body} index
of \vertex{3}.
This results in the bottom graph of \autoref{fig:Direct Conflicts}.
The conflict between Alice's and Bob's changes is
represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{body}
index of \vertex{3}).
In the graph representation, we flag this conflict by highlighting its edges
in red.
In the textual representation, we flag this conflict with the notation~\texttt{\{ 2 | f x \}}.
This is a conflict that much eventually be resolved by the user by deleting
one of the edges and editing the other to be the merged value.
For example, one could delete \edge{10} and then edit the~\texttt{x}
to be~\texttt{x + 2}.
Note that users can continue to edit the code while in this conflicted state.
It just must be resolved before the the code can be run.\footnote{
It may be possible to develop evaluation models that allow these sorts of conflicts,
but that is beyond the scope of this paper.}

@@

TODO TOP: +(\hole, +(\hole, x))

TODO LEFT: +(x, +(\hole, \hole))

TODO RIGHT: +(\hole, +(x, \hole))

TODO BOTTOM: Two parents pointing to x

\caption{Multi-parent conflict}
\label{fig:Multi-parent conflict}
\end{figure}

An example of this is shown in \autoref{fig:Multi-parent conflict},
which starts with the top graph, which represents the expression~\texttt{\hole + (\hole + x)}.
Suppose Alice moves the variable~\texttt{x} to the left-most hole.
This results in the left graph, which represents the expression~\texttt{x + (\hole + \hole)}.
At the same time, suppose Bob moves the variable~\texttt{x} to the right-most hole.
This results in the right graph, which represents the expression~\texttt{\hole + (x + \hole)}.
In both cases, these edits are achieved with one edge delete and one edge add.
Alice deletes edge TODO and adds edge TODO, while Bob deletes edge TODO and adds edge TODO.
As explained in Subsection~TODO:REF, deletes are idempotent so both Alice and Bob deleting
edge TODO is not a problem.
However, both edge TODO (added by Alice) and edge TODO (added by Bob) point to the same vertex.
This generates the bottom graph, in which vertex TODO as two parents pointing to it.
We call this situation a multi-parent conflict.

>>>>

\tikzset{graphs/graph style/.style={
 tree layout,
 level sep=1cm,
 nodes={draw, rounded corners=1mm},
 edges={draw, sloped, font={\small\ttfamily}}
}}
\tikzset{graphs/root style/.style={
 %draw=none,
 as={\textbullet$_{0}$}
}}
%\tikzset{graphs/every edge/.style={}

@@

TODO: "run through several senarios though in practice these occur less frequently than portrayed here"

@@

TODO: group user-level edits (with user visible)

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
TODO
\caption{User interface}
\label{fig:A simple tree with a hole-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
   {}
  }
 }
};
\end{tikzpicture}
\caption{Graph representation}
\label{fig:A simple tree with a hole-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{A simple tree with a hole}
\label{fig:A simple tree with a hole}
\end{figure}

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
TODO
\caption{User interface}
\label{fig:Addition-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
   "$\texttt{var(y)}_{6}$" [> "right", > "5"']
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:A simple tree with a hole-b} after adding \vertex{6} and \edge{5}}
\label{fig:Addition-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Addition}
\label{fig:Addition}
\end{figure}

@@

TODO: z -> foo (or multi-editor)

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
   {}
  }
 }
};
\path (2cm,-2cm) graph[graph style] {
 y [as={$\texttt{var(y)}_{6}$}]
};
\end{tikzpicture}
\caption{\autoref{fig:Addition-b} after deleting \edge{5}}
\label{fig:Change-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{*}_{2}$" [> "root", > "1"'] -> {
   "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
   "$\texttt{var(z)}_{8}$" [> "right", > "7"']
  }
 }
};
\path (2cm,-2cm) graph[graph style] {
 "$\texttt{var(y)}_{6}$"
};
\end{tikzpicture}
\caption{\autoref{fig:Change-a} after adding \vertex{8} and \edge{7}}
\label{fig:Change-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Change}
\label{fig:Change}
\end{figure}

@@

\begin{figure}
% TODO: naive add that creates a conflict
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
% \path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
% \path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
% \path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
% \path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style]
};
\path (-1cm,-3cm) graph[graph style] {
 "$\texttt{*}_{2}$" -> {
  "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
  "$\texttt{var(z)}_{6}$" [> "right", > "5"']
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Change-b} after deleting \edge{1}}
\label{fig:Wrapping-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"']
 }
};
\path (-1cm,-3cm) graph[graph style] {
 "$\texttt{*}_{2}$" -> {
  "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
  "$\texttt{var(z)}_{6}$" [> "right", > "5"']
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Wrapping-a} after adding \vertex{8} and \edge{7}}
\label{fig:Wrapping-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "9"'] -> {
    "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
    "$\texttt{var(z)}_{6}$" [> "right", > "5"']
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Wrapping-b} after adding \edge{9}}
\label{fig:Wrapping-c}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Wrapping}
\label{fig:Wrapping}
\end{figure}

@@

\begin{figure}
% \centering
% \begin{tikzpicture}
% \path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
% \path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
% \path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"']
 }
};
\path (0cm,-3cm) graph [graph style] {
 "$\texttt{*}_{2}$" -> {
  "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
  "$\texttt{var(z)}_{6}$" [> "right", > "5"']
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Wrapping-c} after deleting \edge{9}}
\label{fig:Move-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   {},
   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
    "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
    "$\texttt{var(z)}_{6}$" [> "right", > "5"']
   }
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Move-a} after adding \edge{11}}
\label{fig:Move-b}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Moving \vertex{2} from the left child to the right child of \vertex{8}}
\label{fig:Move}
\end{figure}

@@

TODO: mention that id numbers are global

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   {},
   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(z)}_{6}$" [> "right", > "5"']
   }
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Move-b} after Alice replaces \vertex{4} with \vertex{12}}
\label{fig:Editing Different Parts of the Code-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   {},
   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
    "$\texttt{var(x)}_{4}$" [> "left", > "3"'],
    "$\texttt{var(v)}_{14}$" [> "right", > "15"']
   }
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Move-b} after Bob replaces \vertex{6} with \vertex{14}}
\label{fig:Editing Different Parts of the Code-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   {},
   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(v)}_{14}$" [> "right", > "15"']
   }
  }
 }
};
\end{tikzpicture}
\caption{The result of merging \autoref{fig:Editing Different Parts of the Code-a} with \autoref{fig:Editing Different Parts of the Code-b}}
\label{fig:Editing Different Parts of the Code-c}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Editing Different Parts of the Code}
\label{fig:Editing Different Parts of the Code}
\end{figure}

% \begin{figure}
% \centering
% TODO: add screenshot of code text to each node
% TODO: remove orphans from diagram
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/1.dot}};
% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2A.dot}};
% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/2B.dot}};
% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-different-parts-of-the-code/3.dot}};
% \end{tikzpicture}
% \caption{Editing Different Parts of the Code}
% \label{fig:Editing Different Parts of the Code}
% \end{figure}

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   {},
   "$\texttt{*}_{2}$" [> "right", > "11"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   }
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Editing Different Parts of the Code-c} after Alice replaces \vertex{14} with \vertex{16}}
\label{fig:Editing Nested Parts of the Code-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(v)}_{14}$" [> "right", > "15"']
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Editing Different Parts of the Code-c} after Bob move \vertex{2} to the left child of \vertex{8}}
\label{fig:Editing Nested Parts of the Code-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}}
\label{fig:Editing Nested Parts of the Code-c}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Editing Nested Parts of the Code}
\label{fig:Editing Nested Parts of the Code}
\end{figure}

% \begin{figure}
% \centering
% TODO: add screenshot of code text to each node
% TODO: remove orphans from diagram
% TODO: annotate edges with delete and restore
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/1.dot}};
% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2A.dot}};
% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/2B.dot}};
% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/editing-nested-parts-of-the-code/3.dot}};
% \end{tikzpicture}
% \caption{Editing Nested Parts of the Code}
% \label{fig:Editing Nested Parts of the Code}
% \end{figure}

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   },
   {
    "$\texttt{var(x)}_{20}$" [> "right", > "21"']
   }
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Alice ... TODO}
\label{fig:Multi-child-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   },
   {
    "$\texttt{app}_{22}$" [> "right", > "23"'] -> {
     "$\texttt{var(f)}_{24}$" [> "fun", > "25"'],
     "$\texttt{var(y)}_{26}$" [> "arg", > "27"']
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Editing Nested Parts of the Code-c} after Bob ... TODO}
\label{fig:Multi-child-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.3\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{*}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(u)}_{12}$" [> "left", > "13"'],
    "$\texttt{var(w)}_{16}$" [> "right", > "17"']
   },
   {
    "$\texttt{var(x)}_{20}$" [> "right", > "21"'],
    "$\texttt{app}_{22}$" [> "right", > "23"'] -> {
     "$\texttt{var(f)}_{24}$" [> "fun", > "25"'],
     "$\texttt{var(y)}_{26}$" [> "arg", > "27"']
    }
   }
  }
 }
};
\end{tikzpicture}
\caption{The result of merging \autoref{fig:Multi-child-a} with \autoref{fig:Multi-child-b}}
\label{fig:Multi-child-c}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Multi-child conflicts}
\label{fig:Multi-child}
\end{figure}

% \begin{figure}
% \centering
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/1.dot}};
% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2A.dot}};
% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/2B.dot}};
% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/direct-conflicts/3.dot}};
% \end{tikzpicture}
% \caption{Direct Conflicts}
% \label{fig:Direct Conflicts}
% \end{figure}

Finally, we address genuine conflicts between edits.
Consider the top graph of \autoref{fig:Multi-child}.
Suppose Alice inserts the number literal~\texttt{2} into the body of the lambda
while Bob inserts the function application~\texttt{f x} also into the body of the lambda.
These results in the left and right graphs of \autoref{fig:Multi-child} respectively.
In particular pay attention to the edges going out of the \texttt{body} index of \vertex{3}.
In the left graph it is \edge{10}, which points to a number literal,
and in the right graph it is \edge{13}, which points to a function application.

We merge these by including \emph{both} edges as edges going out of the \texttt{body} index
of \vertex{3}.
This results in the bottom graph of \autoref{fig:Multi-child}.
The conflict between Alice's and Bob's changes is
represented by the fact that there are multiple edges going out of one index (i.e. the \texttt{body}
index of \vertex{3}).
In the graph representation, we flag this conflict by highlighting its edges
in red.
In the textual representation, we flag this conflict with the notation~\texttt{\{ 2 | f x \}}.
This is a conflict that much eventually be resolved by the user by deleting
one of the edges and editing the other to be the merged value.
For example, one could delete \edge{10} and then edit the~\texttt{x}
to be~\texttt{x + 2}.
Note that users can continue to edit the code while in this conflicted state.
It just must be resolved before the the code can be run.\footnote{
It may be possible to develop evaluation models that allow these sorts of conflicts,
but that is beyond the scope of this paper.}

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    "$\texttt{var(x)}_{2}$" [> "left", > "19"'],
    {}
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{Initial code}
\label{fig:Multi-parent-a}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'],
   "$\texttt{var(x)}_{2}$" [> "left", > "19"']
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Multi-parent-a} after Alice ... TODO}
\label{fig:Multi-parent-b}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "$\texttt{+}_{8}$" [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] -> {
    {},
    "$\texttt{var(x)}_{2}$" [> "right", > "19"']
   },
   {}
  }
 }
};
\end{tikzpicture}
\caption{\autoref{fig:Multi-child-a} after Bob ... TODO}
\label{fig:Multi-parent-c}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}{0.24\textwidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  a/{$\texttt{+}_{8}$} [> "root", > "7"'] -> {
   "$\texttt{+}_{2}$" [> "left", > "19"'] ->["right", "19"'] {
    {},
    b/"$\texttt{var(x)}_{2}$"
   },
   {}
  }
 },
 a ->["right", "27"'] b
};
\end{tikzpicture}
\caption{The result of merging \autoref{fig:Multi-parent-b} with \autoref{fig:Multi-parent-c}}
\label{fig:Multi-parent-d}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Multi-parent conflicts}
\label{fig:Multi-parent}
\end{figure}
